---
title: "Zhou networks AUPR"
author: "Ji Huang"
date: "2023-04-04"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
        toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The paper is: [Meta Gene Regulatory Networks in Maize Highlight Functionally Relevant Regulatory Interactions](https://academic.oup.com/plcell/article/32/5/1377/6118116)

The data was downloaded from: https://conservancy.umn.edu/handle/11299/212030

Outputs: 
1. `02_zma_zhou_networks_AUPR.pdf`
2. `02_zma_zhou_networks_summary.tsv`
3. `02_zma_zhou_networks_AUPR_dumbell.pdf`

## 0. Prep

```{r, message=FALSE}
library(here)
library(tidyverse)
library(precrec)
library(glue)

source(here("src", "functions","ggplot2_image_settings.R"))

source(here("src", "functions", "aupr_functions.R"))

rep_times <- 1000

ntwk_dir <- here("data", "external_data","Zhou_network_2020", "rf_100k")

tf_4paper <- read_tsv(here("result", "DFG", 
                           "TF_withTARGET_4paper_20230505.txt")) %>% 
    pull(geneID)
```

```{r, message=FALSE, collapse=TRUE}

## 1. Load GRN
load_network <- function(ntwk_file ) {
    grn <- read_tsv(ntwk_file) %>% 
        dplyr::rename(weight = score)
    return(grn)
}
```

Load all maize TARGET results.

```{r, message=FALSE}
ptarget_all <- load_zma_target() %>% 
    filter(regulator %in% tf_4paper)
```

## 1. Load networks and plot PR curves.

```{r}
ntwk_names <- list.files(ntwk_dir)
```

```{r}
list_plots <- list()

ntwk_summary <- tibble(
    nid = "rm1", 
    aupr = 0.88, aupr_max = 0.88, aupr_min = 0.88, aupr_mean =0.88,
    roc = 0.88, roc_max = 0.88, roc_min = 0.88,
    pvalue = 0.88
)
```


```{r}
for (i in ntwk_names) {
    
    message("Processing network: ", i)
    
    dfg <- load_network(here(ntwk_dir, i))
    all_regulator <- unique(dfg$regulator)[unique(dfg$regulator) %in%
                                           unique(ptarget_all$regulator)]
    ptarget <- ptarget_all %>% 
        filter(regulator %in% all_regulator)
    
    filter_edge <- ptarget$edge
    
    num_tf_with_TARGET <- length(all_regulator)
    
    message("There are a total of ", num_tf_with_TARGET, " TFs with TARGET data in the network.");
    
    sort(all_regulator)
    
    ntwk_1 <- run_aupr(input_network = dfg, rep_times = rep_times)
    
    ## AUPR
    GRN_aupr <- round(auc(ntwk_1$mmcurves)[2,4], 3)
    rand_max_aupr <- round(auc(ntwk_1$mmcurves)[4,4], 3)
    rand_min_aupr <- round(auc(ntwk_1$mmcurves)[6,4], 3)
    
    mean_aupr <-  round(mean(ntwk_1$rand_ntwk), 3)
    
    ## ROC
    GRN_roc <- round(auc(ntwk_1$mmcurves)[1,4], 3)
    rand_max_roc <- round(auc(ntwk_1$mmcurves)[3,4], 3)
    rand_min_roc <- round(auc(ntwk_1$mmcurves)[5,4], 3)
    
    ntwk_short <- str_remove(i, ".tsv.gz")
    pval <- round((rep_times - sum(auc(ntwk_1$mmcurves)[2,4] > ntwk_1$rand_ntwk))/rep_times, 3)
    ptitle1 <- glue(ntwk_short, " AUPR=", {GRN_aupr})
    ptitle2 <- glue("Rand_max=", {rand_max_aupr}, " ",
                    "Rand_min=", {rand_min_aupr}, " ",
                    "Rand_mean=", {mean_aupr}, "\n",
                    {num_tf_with_TARGET}, " TFs with TARGET data. ", 
                    "Pvalue = ", pval)
    p_aupr <- plot_PRcurve(input_data = ntwk_1$mmcurves) +
        labs(title = ptitle1, subtitle = ptitle2,
             caption = glue("Rep", rep_times))
    
    list_plots[[i]] <- p_aupr
    
    ntwk_summary <- ntwk_summary %>% 
        add_row(nid = ntwk_short, 
                aupr = GRN_aupr, 
                aupr_max = rand_max_aupr, aupr_min = rand_min_aupr,
                aupr_mean = mean_aupr,
                roc = GRN_roc, roc_max = rand_max_roc, roc_min = rand_min_roc,
                pvalue = pval)

}
```

## 2. Save the output and PR curves.

```{r}
pdf(here("result", "other_networks", "02_zma_zhou_networks_AUPR_4paper.pdf"), 
    width = 5, height = 5)
list_plots
dev.off()
```

Save network summary.

```{r}
studies <- readxl::read_excel(here("data", "external_data", 
                                   "Zhou_network_2020", "studies.xlsx"))

ntwk_summary <- ntwk_summary %>% 
    filter(nid != "rm1") %>% 
    left_join(studies, by = "nid")

write_tsv(ntwk_summary, 
          here("result", "other_networks", 
               "02_zma_zhou_networks_summary_4paper.tsv"))
```

## 3. Plot dumbell plot for the AUPR.

Red is the network AUPR, grey is the random_max AUPR.

```{r}
library(ggalt)

ntwk_summary <- read_tsv(here("result", "other_networks", 
               "02_zma_zhou_networks_summary_4paper.tsv"))

## order by mean
ntwk_summary <- ntwk_summary %>% 
    mutate(max_order = aupr - aupr_mean) %>% 
    mutate(nid_name = if_else(
        subid == "z",
        paste0(study, "_", note, "[", sample_size, "]"),
        paste0(study, "_", subid, "[", sample_size, "]")
        )
    )

## Ordering by improvement
ntwk_summary <- ntwk_summary %>% 
        arrange(max_order) %>% 
    mutate(nid_name = factor(nid_name, levels = nid_name))

```

Plot the mean AUPR.
```{r}
p3 <- ggplot(ntwk_summary, aes(y=nid_name, x=aupr, xend=aupr_mean)) + 
  geom_dumbbell(size=1.5, color="#e3e2e1", 
                colour_x = "#e41a1c", colour_xend = "gray56",
                dot_guide=FALSE, show.legend = TRUE) +
  labs(x="AUPR", y=NULL, title="Networks' AUPR") +
  theme_minimal() +
  theme(panel.grid.major.x=element_line(size=0.05)) +
  theme(panel.grid.major.y=element_blank())
p3
```

```{r}
pdf(here("result", "other_networks", 
         "02_zma_zhou_networks_AUPR_dumbell_4paper_meanAUPR.pdf"), 
    width = 6, height = 6)
p3
dev.off()
```
