---
title: "Plot Nmetabolism genes in NxTime sets"
author: "Ji Huang"
date: "2023-03-28"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Prep

```{r, message=FALSE}
library(here)
library(tidyverse)
library(glue)
library(ggforce)
library(ggsci)
library(ClusterGVis)
library(RColorBrewer)
library(viridisLite)
library(DT)

expr_name <- "zmaShoot_Npath_NxT_expr.pdf"
heatmap_name_order <- "zmaShoot_Npath_NxT_log2FCunscaled_geneOrder_heatmap.pdf"

```

## 1. Load data

```{r, message=FALSE}
mts_quant <- read_tsv(here("result", "quant_normalized_expr_all.tsv.gz"))

zma_shoot <- read_tsv(here("result", "DEgenes", "edge", 
                           "zma_NxTgenes_df5_edge.tsv"))

zma_root <- read_tsv(here("result", "DEgenes", "root",
                          "zma_root_NxTgenes_df5_edge_20210501_2294genes.tsv"))

## Nmetabolism genes from `008_Npathway_gene_combine_with_name.R`.
Npath_genes <- read_tsv(here("result", "misc", 
                             "zma_Npathway_genes_with_names_20230419.tsv")) %>% 
    distinct()

Npath_genes <- Npath_genes %>% 
    filter(!pathway_step %in% c("cyanate degradation",
                                "Glycolate Oxidase"))
```

## 2. How many Npathyway genes in NxTime Shoot and Root?

These are possible functional genes in Nitrogen metabolism in maize.

```{r}
Npath_genes_shoot <- Npath_genes %>% 
    filter(sgene %in% zma_shoot$gene) %>% 
    select(pathway_step, sgene, zma_gene_symbol) %>% 
    distinct()

Npath_genes_root <- Npath_genes %>% 
    filter(sgene %in% zma_root$gene) %>% 
    select(pathway_step, sgene, zma_gene_symbol) %>% 
    distinct()

Npath_genes_summary <- Npath_genes %>% 
    select(pathway_step, sgene) %>% distinct() %>% 
    count(pathway_step, name = "zma_all") %>% 
    left_join(
        Npath_genes_shoot %>% 
            select(pathway_step, sgene) %>% distinct() %>% 
            count(pathway_step, name = "zma_Shoot"),
        by = ("pathway_step")
    ) %>% 
    left_join(
        Npath_genes_root %>% 
            select(pathway_step, sgene) %>% distinct() %>% 
            count(pathway_step, name = "zma_Root"),
        by = ("pathway_step")
    ) %>% 
    replace_na(list(zma_Shoot=0, zma_Root=0))
```

Nmetabolism genes summary.

```{r}
Npath_genes_summary %>% DT::datatable()
```

Nmetabolism genes in **Shoot**

```{r}
Npath_genes_shoot %>% DT::datatable()
```

Nmetabolism genes in **Root**

```{r}
Npath_genes_root %>% DT::datatable()
```

## 3. Plot the gene expression line plots.

For Nmetabolism genes that are in the Zma-Shoot NxTime gene list.

Prep input data.
```{r}
## Convert to long format for ggplot2.
mts_quant <- mts_quant %>% filter(geneid %in% Npath_genes_shoot$sgene) %>% 
    rename(`0mS1-noN` = `0mS1`, `0mS2-noN` = `0mS2`, `0mS3-noN` = `0mS3`) %>% 
    pivot_longer(col=-geneid, names_to = "condition", values_to = "quant") %>% 
    separate(condition, into = c("condition", "Ntreatment"), sep="-") %>% 
    separate(condition, sep = "mS", into = c("time", "rep")) %>% 
    mutate(time = as.integer(time))

## In order to let both curves start from the same point (time0).
## I have to duplicate the time0 data with Ntreatment=="N". I believe this is also
## what Kranthi used.

mts_time0 <- mts_quant %>% filter(time==0) %>% mutate(Ntreatment="N")
mts_quant <- bind_rows(mts_quant, mts_time0)
gene <- unique(mts_quant$geneid)

print_gene_numebr <- length(gene)
print_gene_id <- paste0(gene, collapse = ",")
message(glue('There are {print_gene_numebr} genes in the total expression matrix'))

mts_quant <- mts_quant %>% left_join(Npath_genes_shoot, by = c("geneid"="sgene")) %>% 
    mutate(geneid = paste0(geneid, "_", zma_gene_symbol, "_", gsub(" ", "",pathway_step))) %>% 
    arrange(pathway_step, zma_gene_symbol)
```

Plot and save.

```{r}
final_page = ceiling(print_gene_numebr/9) # changed page=i to page=1

start_time <- Sys.time()

pdf(here("result", "DEgenes", "expr_plots", 
         expr_name), 
    width = 14, height = 12)

for (i in 1:final_page) {
    p1 <- ggplot(data= mts_quant, aes(x=time, y=quant, color=Ntreatment))+
        geom_point() +
        facet_wrap_paginate(~ geneid, ncol = 3, nrow = 3, page = i, scales = "free") +
        cowplot::theme_half_open()+
        scale_x_continuous(breaks=c(0,5,10,15,20,30,45,60,90,120)) + 
        scale_y_continuous()+
        xlab("Time (min)") +
        ylab("Normalized expression (UMI)")+
        stat_smooth(method = lm, 
                    formula = y ~ splines::ns(x, 5), 
                    se = FALSE) +
        theme(panel.grid.major = element_line(color = "grey80",
                                              linewidth = 0.2,
                                              linetype = 1))
    print(p1)
}

dev.off()

end_time <- Sys.time()
```

```{r}
message("Saved at ", here("result", "DEgenes", "expr_plots", expr_name))
message("It took ", end_time - start_time)
```

## 4. Plot the log2FC heatmap for Nmetabolism genes

```{r}
## Load the log2FC data.
source(here("src", "functions", "load_clean_zma_shoot_log2FC_data.R"))

## All genes EXCEPT N-metabolism genes
nxt_foldChange_nxt <- nxt_foldChange %>% 
    filter(!geneid %in% Npath_genes_shoot$sgene)

## ONLY N-metabolism genes
nxt_foldChange <- nxt_foldChange %>% 
    filter(geneid %in% Npath_genes_shoot$sgene)

gene <- unique(nxt_foldChange$geneid)
length(unique(gene))
```

The Nmetabolism genes are mostly activated.

```{r, fig.height = 3, fig.width = 5}
# A quick view on the hist plot
pdf(file = here("result", "DEgenes", "heatmap", "zma_shoot_nxt_log2_hist.pdf"),
    width = 4.2, height = 2.8)
line <- par(lwd=0.5)
hist(nxt_foldChange_nxt$log2FC, probability = T, breaks = 100, 
     col = "grey", xlim=c(-6,8), xlab = "log2FC", 
     main = expression(italic("All Zma Shoot NxTime genes")),
     border = "grey30")
hist(nxt_foldChange$log2FC, probability = T, breaks = 100, 
     col = "darkgreen", xlim=c(-6,8), xlab = "log2FC", add=F,
     main = expression(italic("Npathway genes in Zma Shoot")),
     border = "grey30")
dev.off()
```

```{r}
ks.test(nxt_foldChange_nxt$log2FC, # all NxT genes
        nxt_foldChange$log2FC,     # Nmetabolism genes
        alternative = "greater")   # one-side test
```

add two histogram together.
```{r}
tt1 <- tibble(log2FC = nxt_foldChange_nxt$log2FC,
              type = "All NxTime")
tt2 <- tibble(log2FC = nxt_foldChange$log2FC,
              type = "Npathway")

tt3 <- bind_rows(tt1, tt2)

p_hist <- ggplot(tt3, aes(x=log2FC, color=type, fill=type)) + 
    geom_histogram(aes(y=after_stat(density)), alpha=0.2, 
                   position="identity", bins = 60) +
    scale_color_manual(values=c("#999999", "darkgreen")) + 
    scale_fill_manual(values=c("#999999", "darkgreen"))+
    cowplot::theme_cowplot()

ggsave(filename = here("result", "DEgenes", "heatmap", 
                       "zma_shoot_nxt_log2_hist_combined.pdf"),
       p_hist, height = 2, width = 5)
```

Add gene names.

```{r}
fc_wide <- nxt_foldChange %>% 
    pivot_wider(names_from = time, names_prefix = "T",
                values_from = log2FC) %>% 
    select(geneid, T5, T10, T15, T20, T30, T45, T60, T90, T120) %>% 
    drop_na() %>% 
    as.data.frame()

rownames(fc_wide) <- fc_wide$geneid
fc_wide$geneid <- NULL

# Add Nmetabolism gene names. If in Liu et al, uses their names. If not, use the maizeGDB names.
all_Npath_data <- tibble(rowid = which(rownames(fc_wide) %in% Npath_genes_shoot$sgene),
                         geneid = rownames(fc_wide)[which(rownames(fc_wide) %in% Npath_genes_shoot$sgene)]) %>% 
    left_join(Npath_genes_shoot, by = c("geneid"="sgene"))

rownames(fc_wide)[all_Npath_data$rowid] <- all_Npath_data$zma_gene_symbol
markGenes = rownames(fc_wide)[all_Npath_data$rowid]

## num of clusters and heatmap color pallet.
cluster_num <- 5
rdbu <- rev(brewer.pal(9, "PRGn"))
```


```{r}
#getClusters(fc_wide)
cm_zma2 <- clusterData(exp = fc_wide,
                      cluster.method = "kmeans",
                      cluster.num = cluster_num, 
                      scaleData = FALSE)

```

```{r}
tt1 <- Npath_genes_shoot %>% rename(gene = zma_gene_symbol)

cm_zma2$wide.res <- cm_zma2$wide.res %>% 
    left_join(tt1, by = "gene") %>% 
    mutate(cluster_new = case_when(pathway_step %in% c("Ammonia Transporters",
                                                       "Nitrate Transporters") ~ 1,
                                   pathway_step %in% c("Nitrite Reductase",
                                                       "Nitrate Reductase") ~ 2,
                                   pathway_step %in% c("Glutamate Decarboxylase",
                                                       "Glutamate Dehydrogenase",
                                                       "Glutamate Synthase",
                                                       "Glutamine Synthetase") ~ 3,
                                   pathway_step %in% c("Asparaginase",
                                                       "Asparagine Synthetase",
                                                       "Aspartate Aminotransferase") ~ 4,
                                   pathway_step %in% c("Carbamoylphosphate Synthase",
                                                       "Carbonic Anhydrase",
                                                       "Nitrilase") ~ 5)) %>% 
    arrange(cluster_new) %>% 
    select(everything(), -cluster, cluster=cluster_new, -pathway_step, -sgene)


## Update long.res
tt3 <- cm_zma2$wide.res %>% select(gene, cluster)

cm_zma2$long.res <- cm_zma2$long.res %>% left_join(tt3, by = "gene") %>% 
    select(everything(), -cluster.x, cluster=cluster.y) %>% 
    arrange(cluster)
```

```{r}
p1 <- visCluster(object = cm_zma2,
                 cluster.order = 1:5,
                 ht.col.list = list(
                     col_range = c(-2,0,3),
                     col_color = c(rdbu[1],rdbu[5],rdbu[9])
                 ),
                 show_row_dend = F,
                 annnoblock.text = T,
                 markGenes = markGenes,
                 markGenes.side = "right",
                 plot.type = "heatmap",
                 column_names_rot = 45,
                 ctAnno.col = ggsci::pal_nejm()(5),
                 sample.col = rev(viridis(n = 9)))

pdf(file = here("result", "DEgenes", "heatmap", heatmap_name_order),
    width = 4, height = 6)

p1
dev.off()
```

