---
title: "JIT GO explore"
author: "Ji Huang"
date: "2023-01-23"
output: 
    html_document:
        code_folding: show
        toc: true
        toc_float: true
        toc_depth: 2
---

Summarize the GO terms by:
1. GOsummaries plots
2. Heatmap as in Varala et al.

All the GO results were from `topGO` package results, testing on **BP** terms.

Output:
1. GOsummarize_JIT_topGO.pdf
2. JIT_GO_plot_heatmap_topGO.png. *need to Run chunks. Knit doesn't work for `pheatmap()`*.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Prep

```{r, message=FALSE}
library(here)
library(tidyverse)
library(GOsummaries)

source(here("src", "functions", "go_enrich.R"))

output_dir = here("result", "JIT", "JIT_genes", "GO")
```

## 1. GOsummaries plots

```{r loadtopGO, message=FALSE}
go_topGO <- read_tsv(file = here(output_dir, "topGO_zma",
                      "GO_result_all_JIT.tsv")) %>% 
    mutate(query = factor(query, levels = c("j5", "j10", "j15", "j20", 
                                          "j30", "j45", "j60", 
                                          "j90", "j120")))

go_topGO %>% count(query)
```

Save combined GO table for Supp tables. The non-filter one was used for heatmap.
```{r}
go_topGO_supp <- go_topGO %>% filter(Significant > 1) %>% 
    arrange(query) %>% mutate(type = "zma_shoot_JIT")

write_tsv(go_topGO_supp, here("result", "Supp_tables_datasets", 
                              "zma_shoot_JIT_GO.tsv"))
```

Filter topGO results:

```{r}
go_topGO_filter <- go_topGO %>% 
    filter(Significant > 1) %>% 
    dplyr::group_by(query) %>% 
    slice_min(n = 20, order_by = w01Fisher1  ) %>% 
    dplyr::select(Term = Term, Score = w01Fisher1, query = query) %>% 
    dplyr::ungroup()
```

Plot the JIT genes GOsummarize plots.

```{r, message=FALSE, results = FALSE}
pdf(here(output_dir, "topGO_zma",
         "GOsummarize_JIT_topGO.pdf"),
    width = 3, height = 3)
map(unique(go_topGO_filter$query), plot_go_summary_JIT, 
    go_table = go_topGO_filter)
dev.off()
```

## 2. GO heatmap

```{r, message=FALSE}
library(RColorBrewer)
library(pheatmap)
library(UpSetR)
```

Prepare plot data for heatmap.

```{r}
pdat_go_heatmap <- prepare_GO_heatmap_data(topGO_table = go_topGO)
```

Plot and saved the heatmap. 

```{r}
## Prepare color palette.
fivenum(pdat_go_heatmap)

bk1 <- c(seq(0, 
             (max(pdat_go_heatmap, na.rm = TRUE)+0.5), 
             by=0.2))

mycolor <- c(colorRampPalette(brewer.pal(9,"Blues"))(length(bk1)-2))

# Plot
png(filename = here(output_dir, "topGO_zma",
                    "JIT_GO_plot_heatmap_topGO.png"), 
    width = 3, height = 4.5, units = "in", res=600)
p1 <- pheatmap(pdat_go_heatmap, cluster_rows = F, cluster_cols = F, color=mycolor,
         show_rownames=F, breaks = bk1, drop_levels = F,
         main = "Maize JIT GO terms (BP)",
         angle_col=0,
         fontsize_col = 14,
         border_color = "NA")
p1
dev.off()
```
