---
title: "JIT motif enrichment"
author: "Ji Huang"
date: "2023-02-07"
output: 
    html_document:
        code_folding: show
        toc: true
        toc_float: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Prep

```{r, message=FALSE}
library(tidyverse)
library(here)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(UpSetR)

source(here("src", "functions", "motif_JIT.R"))
```

Load the Matt's cluster names.

```{r}
cluster80 <- readxl::read_xlsx(here("data", "external_data", 
                            "ClusterDesc_NCor045.xlsx")) %>% 
    janitor::clean_names() %>% 
    mutate(ID = paste0("cluster_", cluster))
```

## 1. Matt's cluster 80 motifs

The MEME-SEA was done on 2023-01-27 (with order2 bfile). The output was downloaded from HPC `/scratch/cgsb/coruzzi/jh6577/project/maize/mts_201906/result/JIT_genes/individual_time/SEA_result_cluster80`.

```{r}
output_dir <- here("result", "JIT", "JIT_genes", "motif",
                                 "SEA_result_cluster80")
```

```{r}
plot_heatmap_motif(outdir = output_dir,
                   motif_db = "cluster80", qval_cutoff = 0.05, ENR_cutoff = 3)
```

```{r}
f_files <- list.files(output_dir, pattern = "sea.tsv", 
                         recursive = TRUE, full.names = T)
motif_cluster <- load_SEA_output(SEA_output_dir = output_dir) %>% 
    filter(QVALUE < 0.05)  %>% filter(ENR_RATIO > 3) %>% 
    left_join(., cluster80, by = "ID") %>% 
    dplyr::select(-DB, -ALT_ID) %>% 
    dplyr::mutate(type = "zma_shoot_JIT_cluster80")

write_tsv(motif_cluster, here(output_dir, 
                              "zma_shoot_JIT_cluster80_motif.tsv"))
```

## 2. JASPAR 656 motifs

The MEME-SEA was done on 2023-01-27 (with order2 bfile). The output was downloaded from HPC `/scratch/cgsb/coruzzi/jh6577/project/maize/mts_201906/result/JIT_genes/individual_time/SEA_result_JASPA2022`.

```{r}
output_dir <- here("result", "JIT", "JIT_genes", "motif",
                                 "SEA_result_JASPA656")
```

```{r}
plot_heatmap_motif(outdir = ,
                   motif_db = "JASPAR656", qval_cutoff = 0.05, ENR_cutoff = 3)
```

```{r}
f_files <- list.files(output_dir, pattern = "sea.tsv", 
                         recursive = TRUE, full.names = T)
motif_cluster <- load_SEA_output(SEA_output_dir = output_dir) %>% 
            filter(QVALUE < 0.05)  %>% filter(ENR_RATIO > 3) %>% 
    dplyr::select(-DB, -ALT_ID) %>% 
    dplyr::mutate(type = "zma_shoot_JIT_JASPAR656")

write_tsv(motif_cluster, here(output_dir, 
                              "zma_shoot_JIT_JASPAR656_motif.tsv"))
```
